---
title: "STATS - HW 3"
author: "Wenjing Li"
date: "2/20/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

### 1(1)

g1 will be smaller. Since the second term means the smoothness and the first term means the fitness. $g_2$ has better smoothness compared with $g_1$. $g_2$ will do a better job at testing and $g_1$ wil be better at training.

### 1(2)

g2 will be smaller. The reason is above.

### 1(3)

If $\lambda = 0$, then $g_1 = g_2$. And $g_1, g_2$ are equal to the MSE function. The training error will be smaller, and the testing error will be larger than above $g$ that has been selected.
## Question 2

```{r,echo=FALSE, warning=FALSE}
#loads data
dat = read.delim("C:/Users/wenji/Downloads/STATS 503/HW3/ozone_data.txt")

#loads package
suppressMessages(suppressWarnings(library(gam)))
```

### 2(1)

At first, we need to take a look at our data set.

```{r}
# Make a summary for the current data set.
summary(dat)
```

From the above result, all the data has been cleaned and we can fit the model directly.

Next, split the data set into training data and testing data and fit a linear model.

```{r}
# Split data - 70% as training data & 30% as testing data
set.seed(123)
n = dim(dat)[1]
train_id = sample(seq(1, n, 1), floor(n*0.7))
test = dat[-train_id, ]
train = dat[train_id, ]

# Fit linear model based on training data
lm = lm((ozone)^(1/3) ~ ., data = train)
summary(lm)
```
From the linear model result, three predictors are all significant. The final linear model based on the training data is:

$ozone^\frac{1}{3} = -0.852 + 0.002*radiation + 0.058*temperature - 0.066*wind$

The p-value of the model is less than 2.2e-16, which means the model is significant. The R-squared is 0.71 which is relatively large. Overall, the linear model gives us a good fit.


### 2(2)

Next, I fitted a GAM on the training data and got results as below.

```{r, warning=FALSE}
# Fit GAM model
Kfold_CV <- function(K,train,h,j,k) {
  fold_size = floor(nrow(train)/K)
  cv_error = rep(0,K)
  for(i in 1:K) {
    if(i != K) {
      CV_test_id = ((i-1)*fold_size+1):(i*fold_size)
    }else{
      CV_test_id = ((i-1)*fold_size+1):nrow(train)
    }
    CV_train = train[-CV_test_id,]
    CV_test = train[CV_test_id,]
    # Fit gam
    gam_CV = gam((ozone)^(1/3) ~ ns(radiation,h)+ns(temperature,j)+
                   ns(wind,k), data=CV_train)
    pred_CV = predict.Gam(gam_CV, CV_test)
    # Calculate CV error by taking averages
    cv_error[i] = mean((CV_test$ozone - pred_CV)^2)
  }
  return(mean(cv_error))
}


K = 10
mse = rep(0, 2000)
dim(mse) = c(10,10,20)
fold_size = floor(nrow(train)/K)
cv_error = rep(0,K)
for (h in 1:10) {
 for (j in 1:10) {
   for (k in 1:20) {
     mse[h,j,k] = Kfold_CV(K,train,h,j,k)
   }
 }
}

min_error = min(mse)

gam = gam((ozone)^(1/3) ~ ns(radiation,4)+ns(temperature,6)+s(wind,2), data=train )
par(mfrow = c(1,3))
plot.Gam(gam)
```

That seems a good fit to the raw data.

```{r}
par(mfrow = c(1,3))
plot(train$radiation, train$ozone)
plot(train$temperature, train$ozone)
plot(train$wind, train$ozone)
```

### 2(c)

```{r}
#compute mse_gam
pred_gam = predict.Gam(gam, test)
mse_gam = mean((pred_gam - test$ozone)^2)

#compute mse_lm
pred_lm = predict.lm(lm,test)
mse_lm = mean((pred_lm - test$ozone)^2)

#create table
table_error = c(mse_lm, mse_gam)
dim(table_error) = c(1,2)
colnames(table_error) = c("Linear Model", "GAM")
rownames(table_error) = c("MSE")
cap = paste("*Testing and Training Error for LM and GAM*")
knitr::kable(table_error, caption = cap)
```

The MSE of the linear model is 2012.6, is larger than the MSE of the GAM model.

### 2(d)

From the raw data plot, the predictor "radiation" is not linear with the response "ozone". The other two predictors can be seen as linear or non-linear predictors. And From the MSE above, the GAM model provides a better fit and choosing the non-linear relationship is better than the linear relationship.